---
title: "p8105_hw2_WZ2591"
author: "Wenyu Zhang"
date: "10/5/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1
## Read and clean the Mr. Trash Wheel sheet:

1. specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

2. use reasonable variable names

3. omit rows that do not include dumpster-specific data

4. round the number of sports balls to the nearest integer

```{r}
mr_trash_wheel = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  filter(!is.na(dumpster), dumpster != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```

## Read and clean precipitation data for 2018 and 2019:

```{r}
trash_wheel_precipitation_2018 = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") %>%
  janitor::clean_names() %>%
  rename(total_precipitation = total) %>%
  filter(!is.na(total_precipitation)) %>%
  mutate(year = 2018, 
         month = month.name[month]) %>%
  relocate(year)
```

```{r}
trash_wheel_precipitation_2019 = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  rename(total_precipitation = total) %>%
  drop_na(total_precipitation) %>% 
  mutate(year = 2019, 
         month = month.name[month]) %>% 
  relocate(year)
```

```{r}
# Combine precipitation datasets for 2018 and 2019
trash_wheel_precipitation_2018_2019 = 
  bind_rows(trash_wheel_precipitation_2018, trash_wheel_precipitation_2019)

trash_wheel_precipitation_2018_2019
```

## Write a paragraph about these data

For "mr_trash_wheel_data", there are `r nrow(mr_trash_wheel)` rows and `r ncol(mr_trash_wheel)` columns with 14 variables: dumpster, month, year, date, weight, volume, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls and homes_powered.

```{r, include=FALSE}
# Describe mr_trash_wheel data
skimr::skim(mr_trash_wheel)
nrow(mr_trash_wheel)
ncol(mr_trash_wheel)
```

For "trash_wheel_precipitation_2018_2019" data, there are `r nrow(trash_wheel_precipitation_2018_2019)` rows and `r ncol(trash_wheel_precipitation_2018_2019)` columns with 3 variables: month, year and total_precipitation

```{r, include=FALSE}
# Describe trash_wheel_precipitation_2018_2019 data
skimr::skim(trash_wheel_precipitation_2018_2019)
nrow(trash_wheel_precipitation_2018_2019)
ncol(trash_wheel_precipitation_2018_2019)
```

For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?

```{r}
# The total precipitation in 2018
sum(pull(trash_wheel_precipitation_2018,total_precipitation))

# The median number of sports balls in a dumpster in 2019
balls_in_2019 = filter(mr_trash_wheel, year == "2019")
median(pull(balls_in_2019, sports_balls))
```

The total precipitation in 201 is `r sum(pull(trash_wheel_precipitation_2018,total_precipitation))`, and the median number of sports balls in a dumpster in 2019 is `r median(pull(balls_in_2019, sports_balls))`

# Problem 2
## First, clean the data in pols-month.csv. 

1. Use separate() to break up the variable mon into integer variables year, month, and day

2. replace month number with month name

3. create a president variable taking values gop and dem, and remove prez_dem and prez_gop

4. remove the day variable.

```{r, message=FALSE}
pols_month_df = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%   
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.name[as.numeric(month)], 
         year = as.numeric(year)) %>% 
  mutate(president = if_else(prez_gop == 0, "democreatic", "republican")) %>%
  select(-prez_dem, -prez_gop, -day)
```

## Second, clean the data in snp.csv using a similar process to the above. 

For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r, message=FALSE}
snp_df = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  arrange(year, month) %>% 
  mutate(month = month.name[as.numeric(month)], 
         year = as.numeric(year)) %>%
  select(-day) 
```

## Third, tidy the unemployment data so that it can be merged with the previous datasets. 

This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r, message=FALSE}
unemployment_df = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  rename(year = Year) %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  mutate(month = month.name[match(month,month.abb)])
```

## Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
pols_snp = 
  left_join(pols_month_df, snp_df, by = c("year", "month"))
pols_snp_unemployment = 
  left_join(pols_snp, unemployment_df, by = c("year", "month"))
```

## Write a short paragraph about these datasets

```{r, include=FALSE}
# describe pols_month_df data
dim(pols_month_df)
range(pull(pols_month_df, year))
names(pols_month_df)

# describe snp_df data
dim(snp_df)
range(pull(snp_df, year))
names(snp_df)

# describe unemployment_df data
dim(unemployment_df)
range(pull(unemployment_df, year))
names(unemployment_df)

# describe pols_snp_unemployment data
dim(pols_snp_unemployment)
range(pull(pols_snp_unemployment, year))
names(pols_snp_unemployment)
```

For describe pols_month_df data, the dimension of data is `r dim(pols_month_df)`, the range of years is `r range(pull(pols_month_df, year))` and names of key variables are `r names(pols_month_df)`

For describe snp_df data, the dimension of data is `r dim(snp_df)`, the range of years is `r range(pull(snp_df, year))` and names of key variables are `r names(snp_df)`

For describe unemployment_df data, the dimension of data is `r dim(unemployment_df)`, the range of years is `r range(pull(unemployment_df, year))` and names of key variables are `r names(unemployment_df)`

For describe pols_snp_unemployment data, the dimension of data is `r dim(pols_snp_unemployment)`, the range of years is `r range(pull(pols_snp_unemployment, year))` and names of key variables are `r names(pols_snp_unemployment)`

# Problem 3
## Load and tidy the data. 

```{r, message=FALSE}
popular_names = 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  mutate(ethnicity = 
           recode(ethnicity, 
                            "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", 
                            "BLACK NON HISP" = "BLACK NON HISPANIC", 
                            "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(
    childs_first_name = toupper(childs_first_name), 
    gender = toupper(gender), 
    ethnicity = toupper(ethnicity)) %>%
  distinct()
```

## Produce a well-structured, reader-friendly table 

Showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. 

```{r}
rank_olivia = 
  popular_names %>% 
  filter(childs_first_name %in% c("OLIVIA"), gender %in% c("FEMALE")) %>%
  select(-count,-childs_first_name,-gender) %>%
  pivot_wider(
              names_from = year_of_birth, 
              values_from = rank) %>%
  knitr::kable()
```

Produce a similar table showing the most popular name among male children over time.

```{r}
popular_male = 
  popular_names %>%
  filter(gender %in% c("MALE"), rank == "1") %>%
  select(-count,-rank,-gender) %>% 
  pivot_wider(
              names_from = year_of_birth, 
              values_from = childs_first_name) %>%
  knitr::kable()
```

## Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
non_hispanic_2016 = 
  popular_names %>%
  filter(ethnicity %in% c("WHITE NON HISPANIC"), gender %in% c("MALE"), year_of_birth %in% c("2016") )

scatter_plot = 
  non_hispanic_2016 %>% 
  ggplot(aes(x = rank, y = count, color = "blue")) + 
  geom_point() +
  labs(title = "the number of children with a name vs the rank in popularity of that name") 

scatter_plot
```








