---
title: "p8105_hw2_WZ2591"
author: "Wenyu Zhang"
date: "10/5/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1
## Read and clean the Mr. Trash Wheel sheet:

1. specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

2. use reasonable variable names

3. omit rows that do not include dumpster-specific data

4. round the number of sports balls to the nearest integer

```{r}
mr_trash_wheel_data = 
  read_excel("data/Trash_Wheel_Collection_Totals_8_6_19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume_cubic = volume_cubic_yards) %>% 
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

mr_trash_wheel_data
```

## Read and clean precipitation data for 2018 and 2019:

```{r}
trash_wheel_precipitation_2018 = 
  read_excel("data/Trash_Wheel_Collection_Totals_8_6_19.xlsx", sheet = 5, range = "A2:B15") %>%
  janitor::clean_names() %>%
  rename(total_precipitation = total) %>%
  filter(!is.na(total_precipitation)) %>% 
  mutate(year = 2018, 
         month = month.name[month])

trash_wheel_precipitation_2018
```

```{r}
trash_wheel_precipitation_2019 = 
  read_excel("data/Trash_Wheel_Collection_Totals_8_6_19.xlsx", sheet = 4, range = "A2:B15") %>%
  janitor::clean_names() %>%
  rename(total_precipitation = total) %>%
  drop_na(total_precipitation) %>% 
  mutate(year = 2019, 
         month = month.name[month])

trash_wheel_precipitation_2019
```

```{r}
trash_wheel_precipitation_2018_2019 = 
  full_join(trash_wheel_precipitation_2018, trash_wheel_precipitation_2019, by = "month")


trash_wheel_precipitation_2018_2019$month[13] = "Grand Total"

trash_wheel_precipitation_2018_2019
```


# Problem 2
##First, clean the data in pols-month.csv. 

1. Use separate() to break up the variable mon into integer variables year, month, and day

2. replace month number with month name

3. create a president variable taking values gop and dem, and remove prez_dem and prez_gop

4. remove the day variable.

```{r}

```

